% Geometry, font
\documentclass[12pt, letter]{article}
\usepackage[margin=0.8in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{fourier}
\usepackage{titling}
\setlength{\droptitle}{-4em} 
\usepackage[parfill]{parskip}
\usepackage{graphicx}
\usepackage{hyperref}

\usepackage{fancyhdr}
\lhead{ECE 4542: Advanced Engineering Algorithms}
\rhead{Waleed Meleis}
\pagestyle{fancy}

% Math stuff
\usepackage{amssymb}
\usepackage{bm}


\author{Zach Neveu}
\title{ Homework 1 }

\begin{document}
\maketitle
\thispagestyle{fancy}

\begin{enumerate}
	\item Bubble Sort
	\begin{enumerate}
		\item Best-case running time: If the list is already sorted, Bubble Sort only requires a single pass, or $O(n)$ time to verify this.
		\item Worst-case running time: If the list is reverse-sorted, then every item will need to be swapped on each pass, which makes the worst case runtime $O(n^2)$
	\end{enumerate}
	\item Sequential Search
	\begin{enumerate}
		\item Best case running time: Desired item is first in array, leading to a runtime of $O(1)$
		\item Worst case running time: Desired item is last in array, leading to a runtime of $O(n)$
	\end{enumerate}
	\item Proofs
	\begin{enumerate}
		\item \textbf{True}: $f(n) = O(g(n))$ means that $f(n)$ is upper-bounded by $g(n)$. Likewise, $g(n) = O(h(n))$ means that $g(n)$ is upper-bounded by $h(n)$. By the transitive property, $f(n) \le g(n) \le h(n)$, so $f(n) = O(h(n))$.
		\item \textbf{False}: If $f(n)$ is linear search (worst case runtime $O(n)$) and $g(n)$ is bubble sort (worst case runtime $O(n^2)$), then $f(n) = O(g(n))$, but $g(n) \ne O(f(n))$
	\end{enumerate}
	\item Optimization vs. Decision Problems
	\begin{enumerate}
		\item Minimum Spanning Tree in a weighted graph
		\begin{itemize}
			\item Decision: Given a weighted graph, determine whether a spanning tree exists for which the sum of the weights of the included edges is less than $k$.
			\item Optimization: Given a weighted graph, find the spanning tree which minimized the sum of the weights of the included edges.
		\end{itemize}
		\item Maximum Matching in a Graph
		\begin{itemize}
			\item Decision: Given a graph, determine whether a matching exists which contains at least $k$ nodes.
			\item Optimization: Given a graph, find the matching containing the largest number of nodes possible.
		\end{itemize}
		\item Shortest Path
		\begin{itemize}
			\item Decision: Given two vertices in a weighted graph, determine whether a path between them exists for which the sum of the included edge weights is less thank $k$.
			\item Optimization: Given two vertices of a weighted graph, find the path between them which has the smallest sum of included edge weights.
		\end{itemize}
	\end{enumerate}
	\item Definitions
	\begin{enumerate}
		\item Heuristic: An approximate method of solving a problem which is not guaranteed to find an optimal solution
		\item Polynomial-time algorithm: An algorithm whos runtime grows as a polynomial function of the instance size
		\item Intractable Problem: A problem which provably cannot be solved by any polynomial-time algorithm.
		\item Complexity class P: All problems which can be solved by a polynomial-time algorithm.
	\end{enumerate}
\end{enumerate}
\end{document}
